Open archaeology, open source?

### 2024-02-07 analysis/_03-data_methodology.qmd
Fix knit problems

```{r data-oarch_forges}
oarch_forges <- oarch |>
  mutate(none = ifelse(is.na(github) & is.na(gitlab) & is.na(bitbucket) &
                         is.na(launchpad) & is.na(codeberg), "", NA)) |>
  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
               names_to = "repo_host", values_to = "repo_url") |>
  drop_na(repo_url) |>
  count(repo_host) |>
  mutate(p = n / sum(n)) |>
  arrange(repo_host == "none", desc(n)) |>
  mutate(repo_host = recode(repo_host,
                            github = "GitHub",
                            codeberg = "Codeberg",
                            gitlab = "GitLab",
                            bitbucket = "BitBucket",
                            launchpad = "Launchpad",
                            none = "None"))

>>>>

```{r data-oarch-forges}
oarch_forges <- oarch |>
  mutate(none = ifelse(is.na(github) & is.na(gitlab) & is.na(bitbucket) &
                         is.na(launchpad) & is.na(codeberg), "", NA)) |>
  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
               names_to = "repo_host", values_to = "repo_url") |>
  drop_na(repo_url) |>
  count(repo_host) |>
  mutate(p = n / sum(n)) |>
  arrange(repo_host == "none", desc(n)) |>
  mutate(repo_host = recode(repo_host,
                            github = "GitHub",
                            codeberg = "Codeberg",
                            gitlab = "GitLab",
                            bitbucket = "BitBucket",
                            launchpad = "Launchpad",
                            none = "None"))

### 2024-02-07 analysis/_04-open_archaeology.qmd
Fix knit problems

```{r num-p_platform_r}
p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / nrow(oarch)
p_platform_python <- sum(oarch$platform == "Python", na.rm = TRUE) / nrow(oarch)
p_platform_qgis <- sum(oarch$platform == "QGIS", na.rm = TRUE) / nrow(oarch)
```

>>>>

```{r data-platforms}
p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / nrow(oarch)
p_platform_python <- sum(oarch$platform == "Python", na.rm = TRUE) / nrow(oarch)
p_platform_qgis <- sum(oarch$platform == "QGIS", na.rm = TRUE) / nrow(oarch)
```

### 2024-02-07 analysis/_05-collaboration.qmd
Fix knit problems

```{r data-oarch_github}
# Subset of dataset with GitHub data
oarch |> 
  filter(!is.na(gh_repo_info)) ->
  oarch_gh
```

@@

```{r data-oarch_lifespan}
oarch_gh %>% 
  unnest(gh_commits) %>% 
  drop_na(datetime) %>% 
  group_by(item_name) %>%
  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
  summarise(
    category = first(category),
    n_contribs = first(n_contribs),
    n = n(),
    first_commit = min(datetime),
    last_commit = max(datetime),
    lifespan = as.integer(last_commit - first_commit),
    commit_rate = n / lifespan
  ) |>
  mutate(commit_rate = if_else(is.finite(commit_rate), commit_rate, NA)) ->
  oarch_lifespan

>>>>

```{r data-oarch-github}
# Subset of dataset with GitHub data
oarch |> 
  filter(!is.na(gh_repo_info)) ->
  oarch_gh
```

@@

```{r data-oarch-lifespan}
oarch_gh %>% 
  unnest(gh_commits) %>% 
  drop_na(datetime) %>% 
  group_by(item_name) %>%
  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
  summarise(
    category = first(category),
    n_contribs = first(n_contribs),
    n = n(),
    first_commit = min(datetime),
    last_commit = max(datetime),
    lifespan = as.integer(last_commit - first_commit),
    commit_rate = n / lifespan
  ) |>
  mutate(commit_rate = if_else(is.finite(commit_rate), commit_rate, NA)) ->
  oarch_lifespan

### 2024-02-07 analysis/_06-network.qmd
Fix knit problems

```{r data-oarch_contributors}
oarch |>
  unnest(gh_contribs) |>
  drop_na(contributor) |>
  group_by(contributor) |>
  mutate(
    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
  ) |>
  group_by(contributor, repo_owner) |>
  summarise(repos = n(), contribs = sum(contributions), .groups = "drop_last") |>
  pivot_wider(id_cols = c(contributor), names_from = repo_owner, 
              values_from = c(repos, contribs), values_fill = 0) |>
  mutate(
    repos = repos_other + repos_own,
    contribs = contribs_other + contribs_own,
    .after = contributor
  ) ->
  oarch_contributors
```

@@

girafe(ggobj = g_repo_user)
```

@@

girafe(ggobj = g_repo_repo)
```

@@

girafe(ggobj = g_user_user)
```

>>>>

```{r data-oarch-contributors}
oarch |>
  unnest(gh_contribs) |>
  drop_na(contributor) |>
  group_by(contributor) |>
  mutate(
    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
  ) |>
  group_by(contributor, repo_owner) |>
  summarise(repos = n(), contribs = sum(contributions), .groups = "drop_last") |>
  pivot_wider(id_cols = c(contributor), names_from = repo_owner, 
              values_from = c(repos, contribs), values_fill = 0) |>
  mutate(
    repos = repos_other + repos_own,
    contribs = contribs_other + contribs_own,
    .after = contributor
  ) ->
  oarch_contributors
```

@@

if(knitr::is_html_output()) girafe(ggobj = g_repo_user) else g_repo_user
```

@@

if(knitr::is_html_output()) girafe(ggobj = g_repo_repo) else g_repo_user
```

@@

if(knitr::is_html_output()) girafe(ggobj = g_user_user) else g_user_user
```

### 2024-02-07 analysis/openarchaeo_collaboration.qmd
Cache chunk output
Fix knit problems

  In this paper, we investigate modes of collaboration in this emerging community of practice using open-archaeo [open-archaeo.info](open-archaeo.info), a curated list of archaeological software, augmented where available with data on their activity on GitHub.
  We apply network analysis to map the networks of collaboration that have emerged among archaeologists using GitHub, and conduct a quantitative exploratory analysis to characterize the nature and intensity of these collaborations.
  Our results show uneven adoption of open source collaborative practices beyond the basic use of git as a version control system and GitHub to host source code.
  For instance, a majority of repositories have 1–3 contributors, with only a few projects distinguished by an active and diverse developer base.
  Furthermore, community-oriented features, such as GitHub's 'issues' and 'comments', are used in only a minority of repositories.
  Our network analysis similarly demonstrates that archaeological open source is firmly embedded within existing power structures that permeate academic life, and that apparent connections are representative of established professional partnerships.
  As such, we find little evidence to support the notion that archaeologists benefit from the potential positive outcomes that are commonly argued to be the natural results of open source development models – namely, greater degrees of transparency, extensibility and participatory action.
bibliography: references.bib
format:
  html:
    page-layout: article
    toc: true
    number-sections: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r dependencies, include=FALSE}
library("openarchaeoCollaboration") # this package, install with devtools::build()

library("english")
library("fs")
library("ggiraph")
library("ggraph")
library("glue")
library("here")
library("khroma")
library("lubridate")
library("patchwork")
library("rgexf")
library("scales")
library("tidygraph")
library("tidyverse")
library("tinytable")
```

>>>>

  In this paper, we investigate modes of collaboration in this emerging community of practice using open-archaeo [open-archaeo.info](open-archaeo.info), a curated list of archaeological software, augmented where available with data on their activity on GitHub.
  We apply network analysis to map the networks of collaboration that have emerged among archaeologists using GitHub, and conduct a quantitative exploratory analysis to characterize the nature and intensity of these collaborations.
  Our results show uneven adoption of open source collaborative practices beyond the basic use of git as a version control system and GitHub to host source code.
  For instance, a majority of repositories have 1–3 contributors, with only a few projects distinguished by an active and diverse developer base.
  Furthermore, community-oriented features, such as GitHub's 'issues' and 'comments', are used in only a minority of repositories.
  Our network analysis similarly demonstrates that archaeological open source is firmly embedded within existing power structures that permeate academic life, and that apparent connections are representative of established professional partnerships.
  As such, we find little evidence to support the notion that archaeologists benefit from the potential positive outcomes that are commonly argued to be the natural results of open source development models – namely, greater degrees of transparency, extensibility and participatory action.
bibliography: references.bib
format:
  html:
    page-layout: article
    toc: true
    number-sections: true
    embed-resources: true
  pdf:
    documentclass: article
    pagesize: a4
    toc: true
    number-sections: true
    extra_dependencies: [ tabularray ]
    includes:
      in_header: fix-tables.tex # workaround for https://tex.stackexchange.com/questions/611786/misplaced-noalign-because-input-before-booktabs-rule
execute:
  cache: true
  echo: false
---

```{r setup}
options("tinytable_tabularray_placement" = "H")
library("openarchaeoCollaboration") # this package, install with devtools::build()
```

```{r dependencies}
library("english")
library("fs")
library("ggiraph")
library("ggraph")
library("glue")
library("here")
library("khroma")
library("lubridate")
library("patchwork")
library("rgexf")
library("scales")
library("tidygraph")
library("tidyverse")
library("tinytable")
```
